Containers and R---the Rockerverse and beyond

### 2020-01-13 manuscript.Rmd
rewording of RStudio section

@@

@@

>>>>

**RStudio** also maintains a set of [Docker images](https://github.com/rstudio/r-docker) for multiple operating systems including SUSE, Ubuntu, CentOS, and Debian.
The image stacks include an [opinionated R installation](https://github.com/rstudio/r-builds) and ensure the R installation and profile is consistent across the different Linux distributions.
The images also install R in a versioned directory and use a minimal set of build and runtime dependencies,
Furthermore these images are responsible for creating pre-compiled binary R packages for Linux, which supplement the binaries built by CRAN.
These pre-compiled packages dramatically decrease the installation time of R packages on Linux and subsequently can decrease the build time of Docker images \citep{lopp_package_2019}.

@@

A general description of what to consider when creating a Docker image used for data science projects in R is provided on [environments.rstudio.com](https://environments.rstudio.com/docker).

@@

## Using Docker in a Data Science Platform

In addition to embedding RStudio Open Source Server inside of Docker images, such as Rocker, RStudio offers the moduler data science platform [RStudio Team](https://rstudio.com/products/team/), a commercial product allowing teams of data scientists and their respective IT/DevOps groups to develop and deploy code in R and Python.
RStudio team enables these groups to use the benefits of containers, without requiring users to learn new tools or directly interact with containers.
Notable patterns include:

1. _To deploy and manage RStudio Professional Products:_
Many RStudio customers are managing professional products in Docker containers as an alternative to bare-metals servers or virtual machines. Best practices for [running RStudio with Docker containers](https://support.rstudio.com/hc/en-us/articles/360021594513-Running-RStudio-with-Docker-containers) as well as [Docker images](https://github.com/rstudio/rstudio-docker-products) for RStudio's commerical products are available.
1. _To run interactive RStudio sessions on a cluster:_
In the RStudio Server Pro 1.2 release in 2019, RStudio added new functionality called [Launcher](https://solutions.rstudio.com/launcher/overview/).
It gives users the ability to spawn RStudio instances and background jobs in a scalable way on external clusters, e.g., [Kubernetes based on Docker images](https://support.rstudio.com/hc/en-us/articles/360019253393-Using-Docker-images-with-RStudio-Server-Pro-Launcher-and-Kubernetes) or [Slurm](https://slurm.schedmd.com/) clusters, and optionally, with Singularity containers.
A key benefit of Launcher is the ability for R and Python users to interactively work in RStudio without learning about Docker at all - while still leveraging containers and Kubernetes.
Users can submit batch jobs to Kubernetes clusters without writing specific deployment scripts.

[RStudio Connect](https://rstudio.com/products/connect/) is a commercial publishing platform to deploy dashboards, REST APIs, and other R and Python content into production. 
RStudio Connect uses an approach to isolate deployed R applications that is similar to Docker's filesystem and process namespacing and leverages the [RStudio Package Manager](https://rstudio.com/products/package-manager/) to manage R environments including system dependencies.
RStudio Package manager may also be used to [manage their R environments and package versions with Docker](https://environments.rstudio.com/docker) and to ensure deterministic image builds.
